---
title: "Data_Visualization"
author: "Sharon"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

##Loading libraries

```{r}
library("tidyverse")
library("ggplot2")
```


## Using ggplot2

Do cars with big engines use more fuel than cars with small engines?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y=hwy))
```

The plot shows a negative relationship between engine size (displ) and fuel efficiency (hwy). In other words, cars with big engines use more fuel.

## Exercises
1. Run ggplot(data = mpg). What do you see?

```{r}
ggplot(data = mpg)
```
Answer: A coordinate system that you can add layers to.
Answer(T): This code creates an empty plot. The ggplot() function creates the background of the plot, but since no layers were specified with geom function, nothing is drawn.

2. How many rows are in mpg? How many columns?
```{r}
glimpse(mpg)
nrow(mpg)
ncol(mpg)
```
Answer: There are 234 rows and 11 columns in the mpg data frame.

3. What does the drv variable describe? Read the help for ?mpg to find out.
A: The type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

4. Make a scatterplot of hwy vs cyl.

```{r}
ggplot(mpg)+
  geom_point(mapping = aes(cyl, hwy))
```

5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

```{r}
ggplot(mpg)+
  geom_point(mapping = aes(class,drv))
```
A: The resulting scatterplot has only a few points. A scatter plot is not a useful display of these variables since both drv and class are categorical variables. Since categorical variables typically take a small number of values, there are a limited number of unique combinations of (x, y) values that can be displayed. In this data, drv takes 3 values and class takes 7 values, meaning that there are only 21 values that could be plotted on a scatterplot of drv vs. class. In this data, there 12 values of (drv, class) are observed. A simple scatter plot does not show how many observations there are for each (x, y) value. As such, scatterplots work best for plotting a continuous x and a continuous y variable, and when all (x, y) values are unique.

## Aesthetic Mappings

Exercises
1. What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

A: The argument colour = "blue" is included within the mapping argument, and as such, it is treated as an aesthetic, which is a mapping between a variable and a value. In the expression, colour = "blue", "blue" is interpreted as a categorical variable which only takes a single value "blue". If this is confusing, consider how colour = 1:234 and colour = 1 are interpreted by aes().

The following code does produces the expected result.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), colour = "blue")
```

2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

A: The following list contains the categorical variables in mpg: manufacturer, model, trans, drv, fl, class
The following list contains the continuous variables in mpg: displ, year, cyl, cty, hwy

3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x=displ, y=hwy, color=manufacturer))
```

A: When a continuous value is mapped to shape, it gives an error. Though we could split a continuous variable into discrete categories and use a shape aesthetic, this would conceptually not make sense. A numeric variable has an order, but shapes do not. It is clear that smaller points correspond to smaller values, or once the color scale is given, which colors correspond to larger or smaller values. But it is not clear whether a square is greater or less than a circle.
